Learning and comparing functional connectomes between populations

### 2012-12-25 paper.tex
Iter on covariance estimation paragraph

%{{{------ Correlation matrices figure ---------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.66\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_emp_cov_3d_crop}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.327\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov_no_confounds.pdf}%

@@

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation with
different choices of confound regressors -- 
\textbf{Left}: regressing out CompCor signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the default mode network.
% XXX: should I outline the default mode network on the main matrix?
-- \textbf{Upper right}: regressing out only movement parameters. -- 
\textbf{Lower right}:
regressing out movement parameters and global signal mean.
% XXX: should this remark go here?
No frequency filtering was applied here.
\label{fig:correlation_matrices}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

%{{{------ Precision matrices figure -----------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.66\linewidth}%
    \includegraphics[width=\linewidth]{group_l21_prec.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_l21_prec_3d_crop}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.327\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_prec.pdf}%

@@

\caption{Different choice of confound regressors --
\label{fig:icov_regressors}
\textbf{Left}: regressing out only movement parameters --
\textbf{Right}: removal of the global mean,
instead of the white matter, CSF, and CompCor time courses. 
}
\end{figure}
%---------------------------------------------------------------------------}}}

% XXX: in the figures, the inserts must be commented.

% XXX: idea: use the same regions as in the Fransson/Marrelec paper on
% the role of the PCC

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables -- strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on normalized signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
Graph Lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is a good general-purpose approach. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see Fig.\,\ref{fig:icov_estimators}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to set this parameter is via cross-validation
\cite{varoquaux2010c}.

% XXX: should we have a summary paragraph commenting on the structure of
% the graphs and correlation matrices extracted.
% And also commenting on the meaning of regressions and sign of
% correlation coefficients (the whole global mean debate).
% \cite{chai2011} (note: chai2011 used compcor)
% Remark that partial correlations consistently have no negative values.

@@

>>>>

%{{{------ Correlation matrices figure ---------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.666\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_emp_cov_3d_labels}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.33\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov_no_confounds.pdf}%

@@

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation with
different choices of confound regressors -- 
\textbf{Left}: regressing out CompCor signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the .
-- \textbf{Upper right}: regressing out only movement parameters. -- 
\textbf{Lower right}:
regressing out movement parameters and global signal mean.
No frequency filtering was applied here.
\label{fig:correlation_matrices}
When no confounding brain signals are regressed, all regions are heavily
correlated. Regressing out common signal, in the form of well-identified
confounds or a global mean, teases out the structure.
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

%{{{------ Precision matrices figure -----------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.666\linewidth}%
    \includegraphics[width=\linewidth]{group_l21_prec.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_l21_prec_3d_labels}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.33\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_prec.pdf}%

@@

\caption{Inverse covariance matrices for different choice of
confound regressors --
\label{fig:icov_regressors}
\textbf{Left}: regressing out only movement parameters --
\textbf{Right}: removal of the global mean,
instead of the white matter, CSF, and CompCor time courses. 
}
\end{figure}
%---------------------------------------------------------------------------}}}

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables -- strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on normalized signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
Graph Lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is in general a good approach for structure recovery. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see Fig.\,\ref{fig:icov_estimators}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to set this parameter is via cross-validation
\cite{varoquaux2010c}.

% XXX: must comment on the estimator figure

\paragraph{Network structure extracted}
%
The correlation matrices and inverse-covariance matrices that we extract
contain a lot of information on the functional structure of the brain.
First, the correlation matrix (Fig.\,\ref{fig:correlation_matrices})
shows blocks of coactivating regions that can be interpreted as
large-scale functional networks, such as the default mode network. Note
the split in network is not straightforward. Different ordering of the
nodes will reveal different networks. Indeed, because of the presence of
hubs and interleaved networks the picture in terms of segregated networks
is not sufficient to explain full-brain connectivity
\cite{varoquaux2012}. The connectivity matrices, correlation matrix or
inverse-covariance matrix, can be represented as graphs: nodes connected
by weighted edges (inserts Fig.\,\ref{fig:correlation_matrices} and
Fig.\,\ref{fig:icov_estimators}). The inverse-covariance matrix,
capturing partial correlations, appears then as extracting a
\emph{backbone} or \emph{core} of the graph. While such structure have
been used as a way to summarize anatomical brain connectivity graphs
\cite{hagmann2008}, here they have a clear-cut meaning with regards to
the BOLD signal: they give the conditional independence structure between
the regions \cite{varoquaux2012}. In other words, regions $a$ and $b$ are
not connected if the signal that they have in common can be explained by
a third region $c$. In this light, the choice of nuisance regressors to
remove confounding common signal is less critical. Indeed, while with
correlation matrices regressing out the global mean has a drastic
(Fig.\,\ref{fig:correlation_matrices} upper right and lower right), on
inverse covariance it only changes very slightly the resulting matrices
(Fig.\,\ref{fig:icov_regressors}). There have been debates on whether to
regress out certain signals, such as the global mean,  as it induces
negative correlations \cite{murphy2009,chang2009}, and these may seem
surprising: one network is doing the opposite of the other. However,
correlation between two signals only takes its meaning with the
definition of a baseline. A simple picture to explain anti-correlations
between two regions is the presence of a third region, mediating the
interactions. Using this third region as a baseline would amount to
estimate partial correlations in the whole system. Using 
inverse-covariance matrices or partial correlations to understand brain
connectivity makes the interpretation in terms of interactions between
brain regions easier and more robust to the choice of confounds.

% XXX: where to insert this in the above paragraph: cite the
% Fransson/Marrelec paper on the role of the PCC

% \cite{chai2011} (note: chai2011 used compcor)
% Remark that partial correlations consistently have no negative values.

@@

Note here that using a Ledoit-Wolf is probably a good choice for this
application
